---
output:
  slidy_presentation: default
  html_notebook: default
---

## Introduction

The following slides use data sourced from the Mississippi State Department of Health. Data are synthesized by Johns Hopkins University and are available here: [https://github.com/CSSEGISandData/COVID-19]. Code to reproduce these graphs is maintained here: [https://github.com/hunterwyork/ms_covid_viz]. The code can easily be adopted for any state.

These slides are automatically updated. The most recent date of data is visible in the titles of several slides

## Introduction, cont.

All cases are as reported by the MSDH. Cases and case rates do not represent the true number of infected individuals in the state. Rather, they reflect availability of testing, testing usage, and actual infection rates. Better metrics for tracking the progress of the pandemic include death rates, hospitalization rates, and testing positivity rates. Only the former was available easily in this data and is included here. 

All questions can be directed to hunterwyork [at] gmail [dot] com

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = FALSE, message = FALSE, fig.align = "center")
library(dplyr)
library(data.table)
library(readr)
library(ggplot2)
library(sf)
library(tigris)
library(stringr)
library(gganimate)
library(transformr)
library(gifski)
library(ggrepel)
library(magrittr)

files_root <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports_us/"

c.state <- "Mississippi"

# create today variable dependent on the time of day
if (Sys.time() %>% 
    substr(12, 13) %>% 
    as.numeric >= 20){
  today <-  Sys.Date() %>% as.Date()-1} else {
    today <-  Sys.Date() %>% as.Date()-2
  }

#beginning date can go back to April 12
beginning_date <- "2020-06-01" %>% as.Date()

# download death data
"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv" %>%
  url() %>%
  read_csv() %>%
  data.table() -> deaths

deaths[, measure := "deaths"]

# download confirmed case data
"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv" %>%
  url() %>%
  read_csv() %>%
  data.table() -> cases

cases[, measure := "cases"]

# append and reshape
county_data <- rbindlist(list(deaths, cases), fill = T)
county_data_long <- melt(county_data[Province_State == c.state],
                         id.vars = names(county_data)[!names(county_data) %like% "/"] )

# fill in pop for cases data since that' smissing
county_data_long[, Population := mean(Population, na.rm = T),
                 .(FIPS, Province_State, variable)]

# reformat date variable
county_data_long[, date_std := as.Date(variable,format="%m/%d/%y")]

## make sure data variable is in order
county_data_long <- county_data_long[order(date_std)]

# Make a new cases/ new deaths indicator
temp <- county_data_long[measure %in% c("cases", "deaths")]
temp[, value := value - lag(value), by = .(FIPS, measure)]
temp[, measure := paste0("new_", measure)]
county_data_long <- rbind(county_data_long, temp)

# create case and death rates
rates_long <- copy(county_data_long)
rates_long[, value := value/Population]
rates_long[, measure := paste0(measure, "_rate")]

# and append to data
county_data_long <- rbindlist(list(county_data_long, rates_long))

# saveRDS(county_data_long, "C:/Users/hyork/Desktop/covid_maps/county_data_long.rds")
# 
# county_data_long <- read_rds("C:/Users/hyork/Desktop/covid_maps/county_data_long.rds")


#now load in a shapefile
shp <- counties(state = str_sub(rates_long[1,FIPS], 1, -4),
                cb = TRUE)

#saveRDS(shp, "shp.rds")


#shp <- read_rds("C:/Users/hyork/Desktop/covid_maps/shp.rds")

# merge onto county data
shp$GEOID <- shp$GEOID %>% as.numeric()
county_data_long_shp <- merge(county_data_long, shp, 
                              by.x = "FIPS", 
                              by.y = "GEOID") %>% data.table()
```

## Total Laboratory-Confirmed & Reported Cases as of `r today`

```{r, echo = F, message = FALSE}
#preserve max value for plotting
county_data_long_shp[measure == "cases" & date_std == today]$value %>% max() -> max_value
county_data_long_shp[measure == "cases" & date_std == today]$value %>% min() -> min_value

#create scale breaks
c(min_value,100,1000,3000,max_value) -> scale_breaks

# plot
gg1 <- ggplot(county_data_long_shp[measure == "cases" & date_std == today]) + 
  geom_sf(aes(fill = value, geometry = geometry), lwd = .25) + 
  theme_classic() +
  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = scale_breaks, 
                       labels = scale_breaks,
                       limits = c(min_value, max_value))+
  labs(fill = "Total Cases", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = value), 
               color = "white", 
               size = 2)
print(gg1)

```

## Total Deaths Attributable to Covid as of `r today`

```{r, echo = F, message = FALSE}
#preserve max value for plotting
county_data_long_shp[measure == "deaths" & date_std == today]$value %>% max() -> max_value
county_data_long_shp[measure == "deaths" & date_std == today]$value %>% min() -> min_value

#create scale breaks
c(min_value, 2^(1:5), max_value) -> scale_breaks

# plot
gg2 <- ggplot(county_data_long_shp[measure == "deaths" & date_std == today]) + 
  geom_sf(aes(fill = value, geometry = geometry), lwd = .25) + 
  theme_classic() +
  scale_fill_viridis_c(option = "A",
                       trans = scales::pseudo_log_trans(sigma = 1, base = 2), breaks = scale_breaks, 
                       labels = scale_breaks,
                       limits = c(min_value, max_value))+
  labs(fill = "Total Deaths", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = value), 
               color = "green", 
               size = 2)
print(gg2)

```

## Total Laboratory-Confirmed Cases of Covid per 1,000 residents as of `r today`

```{r, echo = F, message = FALSE}
#preserve max value for plotting
county_data_long_shp[measure == "cases_rate" & date_std == today]$value %>% max() -> max_value
county_data_long_shp[measure == "cases_rate" & date_std == today]$value %>% min() -> min_value

#create scale breaks
round(c(seq(0, .04, .005), max_value)*1000, 3) -> scale_breaks

# plot
gg4 <- ggplot(county_data_long_shp[measure == "cases_rate" & date_std == today]) + 
  geom_sf(aes(fill = value*1000, geometry = geometry), lwd = .25) + 
  theme_classic() +
  scale_fill_viridis_c(breaks = scale_breaks,
                       trans = scales::pseudo_log_trans(sigma = .1, base = (2)),
                       labels = scale_breaks,
                       limits = c(min_value*1000, max_value*1000))+
  labs(fill = "Cases/1,000", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = round(value*1000,2)), 
               color = "white", 
               size = 2)
print(gg4)
```



## Total Deaths Attributable to Covid per 1,000 residents as of `r today`

```{r, echo = F, message = FALSE}
#preserve max value for plotting
county_data_long_shp[measure == "deaths_rate" & date_std == today]$value %>% max() -> max_value
county_data_long_shp[measure == "deaths_rate" & date_std == today]$value %>% min() -> min_value

#create scale breaks
round(c(seq(0, .002, .0005), max_value)*1000, 3) -> scale_breaks

# plot
gg3 <- ggplot(county_data_long_shp[measure == "deaths_rate" & date_std == today]) + 
  geom_sf(aes(fill = value*1000, geometry = geometry), lwd = .25) + 
  theme_classic() +
  scale_fill_viridis_c(option = "A",
                       breaks = scale_breaks,
                       trans = scales::pseudo_log_trans(sigma = .1, base = (2)),
                       labels = scale_breaks,
                       limits = c(min_value*1000, max_value*1000))+
  labs(fill = "Deaths/1,000", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = round(value*1000,2)), 
               color = "green", 
               size = 2)
print(gg3)
```

## Regional Trends: Introduction

```{r, echo = F, message = FALSE}


## Assign north, central and southern MS variables
county_data_long <- county_data_long[Lat != 0]
county_data_long[Lat <= min(Lat) + ((max(Lat) - min(Lat))/3) & Lat != 0, region := "South"]
county_data_long[Lat > min(Lat) + ((max(Lat) - min(Lat))/3) & Lat <= min(Lat) + ((max(Lat) - min(Lat))/1.5) & Lat != 0, region := "Central "]
county_data_long[Lat > min(Lat) + ((max(Lat) - min(Lat))/1.5) & Lat != 0, region := "North"]

county_data_long[Long_ <= min(Long_) + ((max(Long_) - min(Long_))/2) & Long_ != 0, region2 := "western"]
county_data_long[Long_ > min(Long_) + ((max(Long_) - min(Long_))/2) & Long_ != 0, region2 := "eastern"]

#map counties by region
merge(unique(county_data_long[,.(region, region2, FIPS)]) , shp, 
                              by.x = "FIPS", 
                              by.y = "GEOID") %>% data.table()  -> regions_county
regions_county[, region_cat := paste0(region, region2)]

ggplot(regions_county) + 
  geom_sf(aes(fill = region_cat, geometry = geometry)) + 
  labs(title = "Regions As Defined For the Purposes of This Analysis",
       legend = "Region") + 
  theme_void()
```

## Regional Trends in New Covid Incident Cases/Day

```{r, echo = F, message = FALSE}

#now make logged variables
for (lag_day in 1:6){
  county_data_long[, paste0("lag_", lag_day, "_value") := data.table::shift(value, lag_day, type = "lag"), by = .(FIPS, measure)]
}

#now make weighted value of past week
county_data_long[, mean_lag_value := (value + lag_1_value + lag_2_value + lag_3_value + lag_4_value + lag_5_value + lag_6_value )/ 7, by = .(FIPS, measure, date_std)]

#collapse to state and region
county_data_long_state <- county_data_long[,.(mean_lag_value = sum(mean_lag_value), value = sum(value), population = sum(Population)), by =.(date_std, measure)]

county_data_long_region <- county_data_long[,.(mean_lag_value = sum(mean_lag_value), value = sum(value), population = sum(Population)), by =.(date_std, measure, region, region2)]

county_data_long_region[, region_cat := paste0(region, "", region2)]
county_data_long_region <- rbind(county_data_long_region, county_data_long_state, fill = T)
county_data_long_region[is.na(region_cat), region_cat := "State"]
county_data_long_region[region_cat == "State", state := "State"]
county_data_long_region[region_cat != "State", state := "Regional"]

ggplot(county_data_long_region[measure == "new_cases" & date_std >= "2020-04-01"  & date_std %in% c(today, seq.Date(as.Date("2020-04-01"), today, 1))]) + geom_line( aes(x = date_std, y = mean_lag_value, group = region_cat, color = region_cat)) +
      theme_bw()+
     # ylim(0, 0.042*1000) + 
      labs(x = "Date", y = "New Reported Laboratory-Confirmed Cases/Day", title = "Regional Overview of the Daily Incidence of New Cases\nOnly Considering Reported and Laboratory-Confirmed Cases\nActual Cases Are Presumed to be Exponentially Higher")+
      theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
      geom_text_repel(data = county_data_long_region[measure == "new_cases" & date_std == today], aes(x = date_std, y = mean_lag_value, label =region_cat, group = region_cat, color = region_cat), nudge_x = 25) + 
  scale_x_date(limits = c(as.Date("2020-04-01"), today + 50))+

    facet_wrap(~state, scales = "free_y", nrow = 2)
  


```

## Regional Trends in Covid Deaths/Day Over Time

```{r}
ggplot(county_data_long_region[measure == "new_deaths" & date_std >= "2020-04-01"  & date_std %in% c(today, seq.Date(as.Date("2020-04-01"), today, 1))]) + geom_line( aes(x = date_std, y = mean_lag_value, group = region_cat, color = region_cat)) +
    theme_bw()+
    # ylim(0, 0.042*1000) + 
    labs(x = "Date", y = "Deaths Per Day", title = "")+
    theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
    geom_text_repel(data = county_data_long_region[measure == "new_deaths" & date_std == today], aes(x = date_std, y = mean_lag_value, label =region_cat, group = region_cat, color = region_cat), nudge_x = 25) +
    xlim(as.Date("2020-04-01"), today + 50)+
    facet_wrap(~state, scales = "free_y", nrow = 2) #+ 
  # scale_x_continuous(labels = c(today, seq.Date(as.Date("2020-04-01"), today, 10), rep("", 10)), breaks = c(today, seq.Date(as.Date("2020-04-01"), today+100, 10)))

```

```{r}
#make sure variable is in order

n <- 0
gg_traj_list <- list()
for(i in unique(county_data_long$region)){
  for(ii in unique(county_data_long$region2)){
    n <- n + 1
    gg_traj_list[[n]] <- ggplot(county_data_long[measure == "new_cases" & date_std >= "2020-04-01" & region == i & region2 == ii & date_std %in% c(today, seq.Date(as.Date("2020-04-01"), today, 5))]) + 
      geom_line(aes(x = date_std, y = mean_lag_value, group = Admin2, color = Admin2)) +
      theme_bw()+
     # ylim(0, 0.042*1000) + 
          xlim(as.Date("2020-04-01"), today + 50)+
      labs(x = "Date", y = "Confirmed Cases (7-day Average)", title = paste0(i,ii," ", c.state))+
      theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
      geom_text_repel(data = county_data_long[measure == "new_cases" & date_std == max(date_std) & region == i & region2 == ii], aes(x = date_std, y = mean_lag_value, label = Admin2, color = Admin2), nudge_x = 25)
  }
}

```

## Regional Case Trends

```{r, echo = F}
print(gg_traj_list[[1]])
```

## Regional Case Trends

```{r, echo = F}
print(gg_traj_list[[2]])
```

## Regional Case Trends

```{r, echo = F}
print(gg_traj_list[[3]])
```

## Regional Case Trends

```{r, echo = F}
print(gg_traj_list[[4]])
```

## Regional Case Trends

```{r, echo = F}
print(gg_traj_list[[5]])
```

## Regional Case Trends

```{r, echo = F}
print(gg_traj_list[[6]])
```

```{r}
#make sure variable is in order

n <- 0
gg_traj_list <- list()
for(i in unique(county_data_long$region)){
  for(ii in unique(county_data_long$region2)){
    n <- n + 1
    gg_traj_list[[n]] <- ggplot(county_data_long[measure == "new_cases_rate" & date_std >= "2020-04-01" & region == i & region2 == ii & date_std %in% c(today, seq.Date(as.Date("2020-04-01"), today, 5))]) + 
      geom_line(aes(x = date_std, y = mean_lag_value, group = Admin2, color = Admin2)) +
      theme_bw()+
     # ylim(0, 0.042*1000) + 
          xlim(as.Date("2020-04-01"), today + 50)+
      labs(x = "Date", y = "Confirmed Cases/1,000 People (7-day Average)", title = paste0(i,ii," ", c.state))+
      theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
      geom_text_repel(data = county_data_long[measure == "new_cases_rate" & date_std == max(date_std) & region == i & region2 == ii], aes(x = date_std, y = mean_lag_value, label = Admin2, color = Admin2), nudge_x = 25)
  }
}

```

## Regional Case Rate Trends

```{r, echo = F}
print(gg_traj_list[[1]])
```

## Regional Case Rate Trends

```{r, echo = F}
print(gg_traj_list[[2]])
```

## Regional Case Rate Trends

```{r, echo = F}
print(gg_traj_list[[3]])
```

## Regional Case Rate Trends

```{r, echo = F}
print(gg_traj_list[[4]])
```

## Regional Case Rate Trends

```{r, echo = F}
print(gg_traj_list[[5]])
```

## Regional Case Rate Trends

```{r, echo = F}
print(gg_traj_list[[6]])
```

```{r}
#make sure variable is in order

n <- 0
gg_traj_list <- list()
for(i in unique(county_data_long$region)){
  for(ii in unique(county_data_long$region2)){
    n <- n + 1
    gg_traj_list[[n]] <- ggplot(county_data_long[measure == "deaths" & date_std >= "2020-04-01" & region == i & region2 == ii & date_std %in% c(today, seq.Date(as.Date("2020-04-01"), today, 5))]) + 
      geom_line(aes(x = date_std, y = value, group = Admin2, color = Admin2)) +
      theme_bw()+
     # ylim(0, 0.042*1000) + 
          xlim(as.Date("2020-04-01"), today + 50)+
      labs(x = "Date", y = "Total Confirmed Deaths", title = paste0(i,ii," ", c.state))+
      theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
      geom_text_repel(data = county_data_long[measure == "deaths" & date_std == max(date_std) & region == i & region2 == ii], aes(x = date_std, y = value, label = Admin2, color = Admin2), nudge_x = 25)
  }
}

```

## Regional Death Trends

```{r, echo = F}
print(gg_traj_list[[1]])
```

## Regional Death Trends

```{r, echo = F}
print(gg_traj_list[[2]])
```

## Regional Death Trends

```{r, echo = F}
print(gg_traj_list[[3]])
```

## Regional Death Trends

```{r, echo = F}
print(gg_traj_list[[4]])
```

## Regional Death Trends

```{r, echo = F}
print(gg_traj_list[[5]])
```

## Regional Death Trends

```{r, echo = F}
print(gg_traj_list[[6]])
```

```{r}
#make sure variable is in order

n <- 0
gg_traj_list <- list()
for(i in unique(county_data_long$region)){
  for(ii in unique(county_data_long$region2)){
    n <- n + 1
    gg_traj_list[[n]] <- ggplot(county_data_long[measure == "deaths_rate" & date_std >= "2020-04-01" & region == i & region2 == ii & date_std %in% c(today, seq.Date(as.Date("2020-04-01"), today, 5))]) + 
      geom_line(aes(x = date_std, y = value*1000, group = Admin2, color = Admin2)) +
      theme_bw()+
     # ylim(0, 0.042*1000) + 
          xlim(as.Date("2020-04-01"), today + 50)+
      labs(x = "Date", y = "Total Confirmed Deaths/1,000 People", title = paste0(i,ii," ", c.state))+
      theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
      geom_text_repel(data = county_data_long[measure == "deaths_rate" & date_std == max(date_std) & region == i & region2 == ii], aes(x = date_std, y = value*1000, label = Admin2, color = Admin2), nudge_x = 25)
  }
}

```

## Regional Death Rate Trends

```{r, echo = F}
print(gg_traj_list[[1]])
```

## Regional Death Rate Trends

```{r, echo = F}
print(gg_traj_list[[2]])
```

## Regional Death Rate Trends

```{r, echo = F}
print(gg_traj_list[[3]])
```

## Regional Death Rate Trends

```{r, echo = F}
print(gg_traj_list[[4]])
```

## Regional Death Rate Trends

```{r, echo = F}
print(gg_traj_list[[5]])
```

## Regional Death Rate Trends

```{r, echo = F}
print(gg_traj_list[[6]])
```

## Development of New Laboratory-Confirmed Cases Over the Past 2 Months

```{r,echo = F, message = FALSE}
county_data_long_shp <- merge(county_data_long, shp, 
                              by.x = "FIPS", 
                              by.y = "GEOID") %>% data.table()

#preserve max value for plotting
county_data_long_shp[measure == "new_cases"]$mean_lag_value %>% max() -> max_value
county_data_long_shp[measure == "new_cases"]$mean_lag_value %>% min() -> min_value

# plot
gg1anim <- ggplot(county_data_long_shp[measure == "new_cases" & date_std %in% seq.Date(today-60,today, 3) ]) +
  geom_sf(aes(fill = mean_lag_value, geometry = geometry), lwd = .25) +
  theme_classic() +
  scale_fill_viridis_c(trans = scales::pseudo_log_trans(),
                       limits = c(min_value, max_value))+
  labs(fill = "7-Day Average Laboratory-Confirmed Cases/Day", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = round(mean_lag_value)),
               color = "white",
               size = 2) +
  labs(title = 'Date: {closest_state}')+
  transition_states(date_std,
                    transition_length = 0,
                    state_length = 1) 
animate(gg1anim, end_pause = 20, duration = 20)


```


## Development of New Laboratory-Confirmed Cases/1000 People Over the Past 2 Months

```{r,echo = F, message = FALSE}
county_data_long_shp <- merge(county_data_long, shp, 
                              by.x = "FIPS", 
                              by.y = "GEOID") %>% data.table()

#preserve max value for plotting
county_data_long_shp[measure == "new_cases_rate"]$mean_lag_value %>% max(., na.rm = T) -> max_value
county_data_long_shp[measure == "new_cases_rate"]$mean_lag_value %>% min(., na.rm = T) -> min_value
min_value <- min_value * 1000
max_value <- max_value * 1000

# plot
gg1anim <- ggplot(county_data_long_shp[measure == "new_cases_rate" & date_std %in% seq.Date(today-60,today, 3) ]) +
  geom_sf(aes(fill = mean_lag_value*1000, geometry = geometry), lwd = .25) +
  theme_classic() +
  scale_fill_viridis_c(trans = scales::pseudo_log_trans(),
                       limits = c(min_value, max_value))+
  labs(fill = "7-Day Average Laboratory-Confirmed Cases/Day/1000 People", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = round(mean_lag_value*1000,3)),
               color = "white",
               size = 2) +
  labs(title = 'Date: {closest_state}')+
  transition_states(date_std,
                    transition_length = 0,
                    state_length = 1) 
animate(gg1anim, end_pause = 20, duration = 20)


```

## Development of Covid-Attributable Over the Past 2 Months

```{r,echo = F, message = FALSE}
county_data_long_shp <- merge(county_data_long, shp, 
                              by.x = "FIPS", 
                              by.y = "GEOID") %>% data.table()

#preserve max value for plotting
county_data_long_shp[measure == "deaths"]$value %>% max() -> max_value
county_data_long_shp[measure == "deaths"]$value %>% min() -> min_value

# plot
gg1anim <- ggplot(county_data_long_shp[measure == "deaths" & date_std %in% seq.Date(today-60,today, 3) ]) +
  geom_sf(aes(fill = mean_lag_value, geometry = geometry), lwd = .25) +
  theme_classic() +
  scale_fill_viridis_c(trans = scales::pseudo_log_trans(),
                       limits = c(min_value, max_value),
                       option = "A")+
  labs(fill = "Cumulative Deaths", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = round(mean_lag_value)),
               color = "green",
               size = 2) +
  labs(title = 'Date: {closest_state}')+
  transition_states(date_std,
                    transition_length = 0,
                    state_length = 1) 
animate(gg1anim, end_pause = 20, duration = 20)


```


## Development of Covid-Attributable Deaths/1000 People Over the Past 2 Months

```{r,echo = F, message = FALSE}
county_data_long_shp <- merge(county_data_long, shp, 
                              by.x = "FIPS", 
                              by.y = "GEOID") %>% data.table()

#preserve max value for plotting
county_data_long_shp[measure == "deaths_rate"]$value %>% max(., na.rm = T) -> max_value
county_data_long_shp[measure == "deaths_rate"]$value %>% min(., na.rm = T) -> min_value
min_value <- min_value * 1000
max_value <- max_value * 1000

# plot
gg1anim <- ggplot(county_data_long_shp[measure == "deaths_rate" & date_std %in% seq.Date(today-60,today, 3) ]) +
  geom_sf(aes(fill = mean_lag_value*1000, geometry = geometry), lwd = .25) +
  theme_classic() +
  scale_fill_viridis_c(trans = scales::pseudo_log_trans(),
                       limits = c(min_value, max_value),
                       option = "A")+
  labs(fill = "Cumulative Deaths/1000 People", x= "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.key.height = unit(3, "line")) +
  geom_sf_text(aes(geometry = geometry, label = round(mean_lag_value*1000,3)),
               color = "green",
               size = 2) +
  labs(title = 'Date: {closest_state}')+
  transition_states(date_std,
                    transition_length = 0,
                    state_length = 1) 
animate(gg1anim, end_pause = 20, duration = 20)


```

## Graphs by County

```{r echo = FALSE, results ='asis'}
county_data_long[measure == "new_cases", measure_neat := "New Laboratory-Confirmed Cases Per Day\nLagged 7-Day Average"]
county_data_long[measure == "cases", measure_neat := "Total Laboratory-Confirmed Cases"]
county_data_long[measure == "new_deaths", measure_neat := "New Covid-Attributable Deaths Per Day\nLagged 7-Day Average"]
county_data_long[measure == "deaths", measure_neat := "Total Covid-Attributable Deaths"]
county_data_long[measure %in% c("deaths", "cases"), value_graph := value]
county_data_long[!measure %in% c("deaths", "cases"), value_graph := mean_lag_value]

for(c.county in unique(county_data_long$Admin2)){
  cat('\n## County Diagnostic Plot: ', c.county, '\n')
  ggg <- ggplot(county_data_long[Admin2 == c.county & date_std %in% seq.Date(today-60,today, 1) & measure %in% c("new_cases", "new_deaths", "cases", "deaths")]) + geom_line(aes(x = date_std, y = value_graph)) + facet_wrap(~measure_neat, scales = "free_y") +
    geom_hline(yintercept = 0, linetype = "dotted")+
    theme_bw() + labs(x = "Date", y = "Value")
  print(ggg)
  cat('\n')
}